---
title: "Evaluating the Jimmy Butlers Fit with Minnesota"
author: "Derek Corcoran"
date: "June 26, 2017"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

## Comparing shot charts from the NBA and Minnesota for 2017

Everyone is very excited about the trade that send Jimmy Butler to the Timberwolves for Zach Lavine, Kris Dunn plus a trade swap. But how does he actually fit's Minnesota's Spatial patterns?

### The Timberwolves were one of the teams that shot threes less frequently

As we see in the shot chart bellow, Minnesota shot threes a lot less frequently than the rest of the league, while attempting more shots than the average NBA team from the elbows.

```{r}
library(SpatialBall)
library(gridExtra)
library(ggplot2)
data(season2017)
```

All this points to a not so modern NBA offense. 


```{r, fig.cap= "Shot chart for the whole NBA compared to Minnesota for 2017"}
NBA <- ShotSeasonGraph(season2017) + ggtitle("NBA shot chart 2017")
Min <- OffShotSeasonGraphTeam(season2017, "Min")
grid.arrange(NBA, Min, ncol = 2)
```

### Projected Starting Linup

Most of the media projects the following Starting Lineup for the Timberwolves

* PG Ricky Rubio
* SG Andrew Wiggins
* SF Jimmy Butler
* PF Karl-Anthony Towns
* C Gorgui Dieng

If we look at the shot chart of that team, we see that none of those players are adept three point shooters

####Rubio

```{r}
OffShotSeasonGraphPlayer(Seasondata = season2017, "Ricky Rubio")

```

####Wiggins

```{r}
OffShotSeasonGraphPlayer(Seasondata = season2017, "Andrew Wiggins")
```

####Butler

```{r}
OffShotSeasonGraphPlayer(Seasondata = season2017, "Jimmy Butler")
```

####Towns

```{r}
OffShotSeasonGraphPlayer(Seasondata = season2017, "Karl-Anthony Towns")
```

####Dieng

```{r}
OffShotSeasonGraphPlayer(Seasondata = season2017, "Gorgui Dieng")
```

```{r, fig.cap= "Shot chart for Minnesota for 2017 taking out Zach LaVine's and Kirs Dunn's shots and adding Jimmy Butler's"}
Miny2018 <- season2017

Miny2018$TEAM_NAME <- ifelse(Miny2018$PLAYER_NAME == "Zach LaVine" | Miny2018$PLAYER_NAME == "Kris Dunn", "Chi", Miny2018$TEAM_NAME)

Miny2018$TEAM_NAME <- ifelse(Miny2018$PLAYER_NAME == "Jimmy Butler", "Min", Miny2018$TEAM_NAME)
Min18 <- OffShotSeasonGraphTeam(Miny2018, "Min") + ggtitle("New Minnesota")
grid.arrange(Min18, Min, ncol = 2)
```