---
title: "Most common and efficient shots per team"
author: "Derek Corcoran"
date: "February 8, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Most common shots in the game
```{r libraries, echo=FALSE, message=FALSE, warning=FALSE}
library(ggplot2)
library(DT)
library(dplyr)
library(knitr)
shotDataTotal2017 <- readRDS("shotDataTotal2017.rds")
```


The nba stats webpage classifies shots as one of the following types *`r unique(shotDataTotal2017$ACTION_TYPE)`*

## Cleaning of the data

This clearly has to be filtered further joined, for example the Layup Shot 2 should be the same category as Driving Layup Shot I think. we can discuss this further.

For now, I found the 6 most tipicall shots made in the NBA, measure by proportion of shots taken overall (Prop). And also calculated the points per shot for each kind of shot(PPS).


```{r structure, echo=FALSE, message=FALSE, warning=FALSE}
shotDataTotal2017$EVENT_TYPE <- as.numeric(as.character(ifelse(shotDataTotal2017$EVENT_TYPE == "Made Shot", 1, 0)))

shotDataTotal2017$SHOT_TYPE <- as.numeric(as.character(ifelse(shotDataTotal2017$SHOT_TYPE == "2PT Field Goal", 2, 3)))

DF <- group_by(shotDataTotal2017, ACTION_TYPE, SHOT_TYPE)
DF <- summarize(DF, Prop = n()/nrow(shotDataTotal2017), prct = sum(EVENT_TYPE)/n())
DF$PPS <- DF$prct*DF$SHOT_TYPE
DF <- dplyr::arrange(DF, desc(Prop))
DF <-head(DF)
DF$ACTION_TYPE <- paste(DF$ACTION_TYPE, DF$SHOT_TYPE)
DF <- DF[,-c(2,4)]
kable(DF, digits = 2)

```


## Sortable table by team

Then I made that information for each team, as you can see, the team that is more monotematic is the Houston Rockets (no surprise there) where more than 40% of their shots are three pointers, you can use the arrows to sort by any of the given categories.Let me know if you think of something with this.



```{r 2, echo=FALSE, message=FALSE, warning=FALSE}

DF2 <- list()
TEAM <- unique(shotDataTotal2017$TEAM_NAME)

for(i in 1:length(TEAM)){
  DF2[[i]]<- dplyr::filter(shotDataTotal2017, TEAM_NAME == TEAM[i])
  shot<- NROW(DF2[[i]])
  DF2[[i]] <- group_by(DF2[[i]], ACTION_TYPE, SHOT_TYPE)
  DF2[[i]] <- summarize(DF2[[i]], Prop = n()/shot, prct = sum(EVENT_TYPE)/n())
  DF2[[i]]$PPS <- DF2[[i]]$prct*DF2[[i]]$SHOT_TYPE
  DF2[[i]] <- dplyr::arrange(DF2[[i]], desc(Prop))
  DF2[[i]]$ACTION_TYPE <- paste(DF2[[i]]$ACTION_TYPE, DF2[[i]]$SHOT_TYPE)
  DF2[[i]]$TEAM <- TEAM[i]
  DF2[[i]] <- filter(DF2[[i]], ACTION_TYPE == "Jump Shot 3" | ACTION_TYPE == "Jump Shot 2" | ACTION_TYPE == "Layup Shot 2" | ACTION_TYPE == "Driving Layup Shot 2" | ACTION_TYPE == "Pullup Jump shot 2" | ACTION_TYPE == "Floating Jump shot 2")
  DF2[[i]] <- DF2[[i]][,-c(2,4)]

}

DF2 <- do.call("rbind", DF2)
DF$TEAM <- c("league")

DF2 <- rbind(DF2, DF)
DF2[,2:3] <- round(DF2[,2:3], digits = 3)
DF2 <- dplyr::arrange(DF2, desc(Prop))


options(DT.options = list(pageLength = 30, language = list(search = 'Filter:')))
DT::datatable(DF2)

```

## Are teams choosing their most efficient shots?

One question worth asking is if teams shoot more the shots that they are more efficient at (more pps).

```{r regr, echo=FALSE, message=FALSE, warning=FALSE}
summary(lm(PPS ~ Prop, data = DF2))

ggplot(DF2, aes(x = Prop, y = PPS)) + geom_point(aes(color = ACTION_TYPE)) + geom_smooth(method = "lm") + theme(legend.position="bottom")

```