---
title: "NBA Finals preview"
author: "Derek Corcoran"
date: "May 27, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

# The long wait

As we wait for the finals


```{r}
library(SpatialBall)
library(knitr)
library(dplyr)
data(season2017)
season2016 <- readRDS("shotDataTotal2016.rds")
season2015 <- readRDS("shotDataTotal2015.rds")
```

```{r, cache=TRUE, results='hide'}
Rating2017<- SpatialRating(Seasondata = season2017)
Rating2015<- SpatialRating(Seasondata = season2015)
Rating2016<- SpatialRating(Seasondata = season2016)

```

```{r}
kable(head(Rating2015,8), caption = "Spatial rating 2015")
```

```{r}
kable(head(Rating2016,8), caption = "Spatial rating 2016")
```

```{r}
kable(head(Rating2017,4), caption = "Spatial rating 2017")
```

```{r}



#Golden State home

Get_Apps(HomeTeam = "GSW", VisitorTeam = "Cle", Seasondata = season2017)

OffShotSeasonGraphTeam(Seasondata = season2017, team = "GSW")
DefShotSeasonGraphTeam(Seasondata = season2017, team = "GSW")

OffShotSeasonGraphTeam(Seasondata = season2017, team = "Cle")
DefShotSeasonGraphTeam(Seasondata = season2017, team = "Cle")


Get_Apps(HomeTeam = "GSW", VisitorTeam = "Cle", Seasondata = season2016)
Get_Apps(HomeTeam = "GSW", VisitorTeam = "Cle", Seasondata = season2015)

#Cleveland Home

Get_Apps(HomeTeam = "Cle", VisitorTeam = "GSW", Seasondata = season2017)
Get_Apps(HomeTeam = "Cle", VisitorTeam = "GSW", Seasondata = season2016)
Get_Apps(HomeTeam = "Cle", VisitorTeam = "GSW", Seasondata = season2015)
```

## Including Plots

You can also embed plots, for example:


This year, at the Bay Area, we predict a spread of `r round(Get_Apps(HomeTeam = "GSW", VisitorTeam = "Cle", Seasondata = season2017)$spread, 2)`, compare that to last year, where before the first game our algorithm would predict a spread of `r round(Get_Apps(HomeTeam = "GSW", VisitorTeam = "Cle", Seasondata = season2016)$spread, 2)` or the first series (2015), where Golden State was predicted 


```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
